### 2022-01-20 manuscript/supplementary_material_validation.Rmd

---
title             : "Keeping an Eye on Hidden Markov Models in Gaze Data Classification: Supplementary Material Validation Study"
shorttitle        : "HMMs in Gaze Classification"

header-includes   :
  - \usepackage{placeins}
floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

bibliography      : "references.bib"
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r packages, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, results = "asis")

library(papaja)
library(tidyverse)
library(here)
library(psych)
library(depmixS4)
library(viridis)
source(here("algorithm/model_helper_functions.R"))
```

```{r set graphics theme, include=FALSE}

theme_set(theme_apa())

```

```{r calculate Schwarz weights for Ehinger et al., include=FALSE}

load(here("validation/Ehinger2019_fitted.Rdata"))

# From Wagenmakers & Farrell (2004)

schwarz.weights <- function(bic, na.rm = T) {

  d.bic <- bic - min(bic, na.rm = na.rm) # eq 2

  exp(-0.5 * d.bic)/sum(exp(-0.5 * d.bic), na.rm = na.rm) # eq 4

}

E2019.bic <- lapply(E2019.fit, function(subj) {
  out <- lapply(subj, function(task) {
    out <- lapply(task, function(block) {
      out <- lapply(1:length(block), function(mod) {

        if(mod == 1) {

          bic <- try(-2*block[[mod]][[1]] + 6*log(block[[mod]][[2]]))

        } else {

          bic <- try(BIC(block[[mod]]$model))

        }

        return(ifelse(is.numeric(bic), bic, NA))
      })

      return(schwarz.weights(unlist(out)))
    })

    df <- as.data.frame(reduce(out, rbind))

    names(df) <- paste("model_", 1:ncol(df), sep = "")

    #blockm <- apply(df, 2, median, na.rm = T)

    return(df)
  })

  return(out)
})

# Create Schwarz weight plots

bic.plot <- list()

bic.plot[[4]] <- reduce(reduce(E2019.bic, rbind)[,1], rbind) %>%
  mutate(Subject = rep(1:15, each = 6),
         Block = rep(1:6, 15)) %>%
  pivot_longer(contains("model"), names_to = "model", values_to = "weight") %>%
  mutate_at(c("Subject", "Block", "model"), as.factor) %>%
  ggplot(aes(x = model, y = Subject)) + geom_tile(aes(fill = weight), color = "black") +
  facet_wrap(vars(Block), ncol = 2, labeller = label_both) +
  scale_x_discrete(name = "Number of states", labels = as.character(1:5)) +
  scale_y_discrete(name = "Subject") +
  scale_fill_viridis_c(name = "Schwarz\nweight", breaks = c(0, 0.2, 0.4, 0.6, 0.8, 1))

bic.plot[[5]] <- reduce(reduce(E2019.bic, rbind)[,2], rbind) %>%
  mutate(Subject = rep(1:15, each = 6),
         Block = rep(1:6, 15)) %>%
  pivot_longer(contains("model"), names_to = "model", values_to = "weight") %>%
  mutate_at(c("Subject", "Block", "model"), as.factor) %>%
  ggplot(aes(x = model, y = Subject)) + geom_tile(aes(fill = weight), color = "black") +
  facet_wrap(vars(Block), ncol = 2, labeller = label_both) +
  scale_x_discrete(name = "Number of states", labels = as.character(1:5)) +
  scale_y_discrete(name = "Subject") +
  scale_fill_viridis_c(name = "Schwarz\nweight", breaks = c(0, 0.2, 0.4, 0.6, 0.8, 1))

```

(ref:plot-schwarz-weights-task4) Schwarz weights displayed for each subject and HMMs with different numbers of states. Models were applied to task 4 of the @Ehinger2019 data set. Higher weights indicate better model fit. Grey tiles indicate erroneous model fits.

```{r plot-schwarz-weights-task4, fig.cap="(ref:plot-schwarz-weights-task4)", fig.height=8}

print(bic.plot[[4]])

```

(ref:plot-schwarz-weights-task5) Schwarz weights displayed for each subject and HMMs with different numbers of states. Models were applied to task 5 of the @Ehinger2019 data set. Higher weights indicate better model fit. Grey tiles indicate erroneous model fits.

```{r plot-schwarz-weights-task5, fig.cap="(ref:plot-schwarz-weights-task5)", fig.height=8}

print(bic.plot[[5]])

```

```{r calculate integrated complete likelihood criterion, include=FALSE}

load(here("validation/Andersson2017_fitted.Rdata"))

# Compute ICL

ICL <- function(mod) {

  -2*logLik(mod, method = "classification") + npar(mod)*log(nobs(mod))

}

A2017.icl <- lapply(A2017.fit, function(stim) {
  out <- lapply(stim, function(subj) {
    out <- lapply(1:length(subj), function(mod) {

      if(mod == 1) {

        icl <- try(-2*subj[[mod]][["LL"]] + 6*log(subj[[mod]][["N"]]))

      } else {

        icl <- try(ICL(subj[[mod]]$model))

      }

      return(ifelse(is.numeric(icl), icl, NA))
    })

    return(schwarz.weights(unlist(out)))
  })

  df <- as.data.frame(reduce(out, rbind))

  names(df) <- paste("model_", 1:length(stim[[1]]), sep = "")

  return(df)
})

# Create ICL weight plots

icl.plot <- lapply(A2017.icl, function(x) {

  data.long <- x %>% 
    mutate(subject = 1:nrow(x)) %>%
    pivot_longer(names(x), names_to = "model", values_to = "weight") %>%
    mutate_at(c("subject", "model"), as.factor)

  p <- ggplot(data.long, aes(x = model, y = subject)) + geom_tile(aes(fill = weight), color = "black") +
    scale_x_discrete(name = "Number of states", labels = as.character(1:5)) +
    scale_y_discrete(name = "Subject") +
    scale_fill_viridis_c(name = "ICL weight", breaks = c(0, 0.2, 0.4, 0.6, 0.8, 1))

  return(p)
})

```

(ref:plot-icl-weights-img) ICL weights displayed for each subject and HMMs with different numbers of states. Models were applied to the image condition of the @Andersson2017 data set. Higher weights indicate better model fit.

```{r plot-icl-weights-img, fig.cap="(ref:plot-icl-weights-img)"}

print(icl.plot[[2]])

```

(ref:plot-icl-weights-dots) ICL weights displayed for each subject and HMMs with different numbers of states. Models were applied to the moving dots condition of the @Andersson2017 data set. Higher weights indicate better model fit.

```{r plot-icl-weights-dots, fig.cap="(ref:plot-icl-weights-dots)"}

print(icl.plot[[1]])

```

(ref:plot-icl-weights-vid) ICL weights displayed for each subject and HMMs with different numbers of states. Models were applied to the video condition of the @Andersson2017 data set. Higher weights indicate better model fit.

```{r plot-icl-weights-vid,  fig.cap="(ref:plot-icl-weights-vid)"}

print(icl.plot[[3]])

```

\FloatBarrier
\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup